---
title: "Homework Assignment 3"
author: "Jing Su"
date: "Due 8am, 2022-12-10"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document:
    includes:
      in_header: ./pset_header.tex
    fig_width: 5
    fig_height: 3.5
---

\begin{small} 

Total points: 400 points

\textbf{Problem set policies.} \textit{Please provide concise, clear answers for each question. Note that only writing the result of a calculation (e.g., "SD = 3.3") without explanation is not sufficient. For problems involving \emph{\textsf{R}}, include the code in your solution, along with any plots.} 
		
\textit{Please submit your problem set via Canvas as a PDF, along with the R Markdown source file.} 
		
\textit{We encourage you to discuss problems with other students (and, of course, with the course head and the TFs), but you must write your final answer in your own words. Solutions prepared "in committee" are not acceptable. If you do collaborate with classmates on a problem, please list your collaborators on your solution.}

\end{small} 


## Unit 4

### Problem 1. (20 points)

In vertebrates, sweet and savory ("umami") tastes are sensed by receptors termed T1Rs. Most vertebrates have three T1Rs, with T1R2 and T1R3 receptors working together to detect sugars (carbohydrates) and artificial sweeteners, while the T1R1-T1R3 heterodimer mediates umami taste. However, even though birds lack \textit{T1R2} genes, several avian species display high behavioral affinity for nectar or sweet fruit. Receptor expression studies in hummingbirds revealed that the ancestral umami receptor (T1R1-T1R3) has been repurposed to detect sugars.\footnote{Baldwin, et al. Evolution of sweet taste perception in hummingbirds by transformation of the ancestral umami receptor. \textit{Science} 2014; \textbf{345}: 929-933.}

Researchers investigated whether T1R1-T1R3 function would dictate hummingbird taste behavior. In a series of field tests, hummingbirds were presented simultaneously with two filled containers, one containing test stimuli and a second containing sucrose. The test stimuli included aspartame, erythritol, water, and sucrose. Aspartame is an artificial sweetener that tastes sweet to humans, but is not detected by hummingbird T1R1-T1R3, while erythritol is an artificial sweetener that is known to activate T1R1-T1R3.

Data on how long a hummingbird drank from a particular container for a given trial, measured in seconds, is in the file \texttt{hummingbirds.Rdata}. Variable names ending in \texttt{1} correspond to the test stimuli, while names ending in \texttt{2} correspond to sucrose. For example, in the first field test comparing aspartame and sucrose, a hummingbird drank from the aspartame container for 0.54 seconds and from the sucrose container for 3.21 seconds.

Do the data suggest that T1R1-T1R3 play the described role in hummingbird taste behavior?

To answer this question, analyze the data for each set of trials: aspartame versus sucrose, erythritol versus sucrose, water versus sucrose, and sucrose versus sucrose. Let $\alpha = 0.05$. Write a conclusion summarizing and interpreting the results, referencing numerical results (such as $p$-values) where appropriate.
```{r}
setwd("C:/Users/VARSHITH REDDY/Downloads")
load('hummingbirds.Rdata')
str(hummingbirds)
```
```{r}
#Summary of  asp.vs.sucr.1, which calculates mean of the sample
summary(hummingbirds$asp.vs.sucr.1)
```
```{r}
#t-test
#We are calculating 95% Confidence interval
t.test(hummingbirds$asp.vs.sucr.1, na.rm = TRUE, conf.level = 0.95)$conf.int
#We now calculate t-statistic , as well as p-value for asp.vs.sucr.1
t.test(hummingbirds$asp.vs.sucr.1, mu = 0.6982300, alternative = 'two.sided')
```
```{r}
#Summary of  asp.vs.sucr.1, which calculates mean of the sample
summary(hummingbirds$asp.vs.sucr.2)

#t-test
#We are calculating 95% Confidence interval
t.test(hummingbirds$asp.vs.sucr.2, na.rm = TRUE, conf.level = 0.95)$conf.int
#We now calculate t-statistic , as well as p-value for asp.vs.sucr.2
t.test(hummingbirds$asp.vs.sucr.1, mu = 2.315, alternative = 'two.sided')
```
```{r}
#Summary of  ery.vs.sucr.1, which calculates mean of the sample
summary(hummingbirds$ery.vs.sucr.1)

#t-test
#We are calculating 95% Confidence interval
t.test(hummingbirds$ery.vs.sucr.1, na.rm = TRUE, conf.level = 0.95)$conf.int
#We now calculate t-statistic , as well as p-value for asp.vs.sucr.1
t.test(hummingbirds$ery.vs.sucr.1, mu = 1.508, alternative = 'two.sided')
```
```{r}
#Summary of  ery.vs.sucr.2, which calculates mean of the sample
summary(hummingbirds$ery.vs.sucr.2)

#t-test
#We are calculating 95% Confidence interval
t.test(hummingbirds$ery.vs.sucr.2, na.rm = TRUE, conf.level = 0.95)$conf.int
#We now calculate t-statistic , as well as p-value for asp.vs.sucr.2
t.test(hummingbirds$ery.vs.sucr.2, mu = 1.487, alternative = 'two.sided')
```
```{r}
#Summary of wat.vs.sucr.1, which calculates mean of the sample
summary(hummingbirds$wat.vs.sucr.1)

#t-test
#We are calculating 95% Confidence interval
t.test(hummingbirds$wat.vs.sucr.1, na.rm = TRUE, conf.level = 0.95)$conf.int
#We now calculate t-statistic , as well as p-value for wat.vs.sucr.1
t.test(hummingbirds$wat.vs.sucr.1, mu = 0.461, alternative = 'two.sided')
```
```{r}
#Summary of wat.vs.sucr.2, which calculates mean of the sample
summary(hummingbirds$wat.vs.sucr.2)

#t-test
#We are calculating 95% Confidence interval
t.test(hummingbirds$wat.vs.sucr.2, na.rm = TRUE, conf.level = 0.95)$conf.int
#We now calculate t-statistic , as well as p-value for wat.vs.sucr.2
t.test(hummingbirds$wat.vs.sucr.2, mu = 2.606, alternative = 'two.sided')
```
```{r}
#Summary of sucr.vs.sucr.1, which calculates mean of the sample
summary(hummingbirds$sucr.vs.sucr.1)

#t-test
#We are calculating 95% Confidence interval
t.test(hummingbirds$sucr.vs.sucr.1, na.rm = TRUE, conf.level = 0.95)$conf.int
#We now calculate t-statistic , as well as p-value for sucr.vs.sucr.1
t.test(hummingbirds$sucr.vs.sucr.1, mu = 1.581, alternative = 'two.sided')
```
```{r}
#Summary of sucr.vs.sucr.2, which calculates mean of the sample
summary(hummingbirds$sucr.vs.sucr.2)

#t-test
#We are calculating 95% Confidence interval
t.test(hummingbirds$sucr.vs.sucr.2, na.rm = TRUE, conf.level = 0.95)$conf.int
#We now calculate t-statistic , as well as p-value for sucr.vs.sucr.2
t.test(hummingbirds$sucr.vs.sucr.2, mu = 1.608, alternative = 'two.sided')
```
#Finding Correlation between each set of variables 
```{r}
cor.test(hummingbirds$asp.vs.sucr.1,hummingbirds$asp.vs.sucr.2)
```
```{r}
cor.test(hummingbirds$ery.vs.sucr.1,hummingbirds$ery.vs.sucr.2)
```
```{r}
cor.test(hummingbirds$wat.vs.sucr.1,hummingbirds$wat.vs.sucr.2)
```
```{r}
cor.test(hummingbirds$sucr.vs.sucr.1,hummingbirds$sucr.vs.sucr.2)
```
#We describe p-value standards as below 0.05 and above 0.05 to reject or accept null hypothesis. Here after performing the correlation and t-test we conclude that asp.sucr and sucr.sucr played a prominent role of a humming bird behaviour. Their pvalues are 0.1876 and 0.3248 respectively. 



### Problem 2. (20 points)

A medical research group is recruiting people to complete short surveys about their medical history. For example, one survey asks about family history of cancer. So far, people who sign up complete an average of 4 surveys, with standard deviation 2.2. The research group wants to try a new interface that they think may encourage new enrollees to complete more surveys. They plan to randomize each enrollee to either the old or new interface. 

  a) How many new enrollees do they need for each group (old or new interface) to detect an effect size of 0.5 surveys per enrollee, if the desired power level is 80\%? Let $\alpha = 0.05$.

```{r}
power.t.test(n = NULL, delta = 4, sd = 2.2, sig.level = 0.05, power = 0.80)
```


  b) Explain the effect of increasing $\alpha$ on the power of the test. What is one disadvantage to increasing $\alpha$, from a decision-making standpoint?
#Usually higher alpha values reject the null hypothesis. If we choose higher alpha values, there is a probability of reducing type 2 errors. The consequence here is that if the null hypothesis is true, increasing alpha makes it more likely that we commit a type 1 error which may lead to severe effects.

## Unit 5

### Problem 3.

Caffeine is the world's most widely used stimulant, with approximately 80\% consumed in the form of coffee. Suppose a study was conducted to investigate the relationship between coffee consumption and exercise. Participants were randomly recruited from the undergraduate and graduate student populations of universities in the Boston/Cambridge area. Participants were asked to report the number of hours they spent per week on moderate (e.g., brisk walking) and vigorous (e.g., strenuous sports and jogging) exercise. Based on these data, the researchers estimated the total hours of metabolic equivalent tasks (MET) per week, a value always greater than 0. The file \texttt{coffee\_exercise.Rdata} contains simulated MET data for the study participants, based on the amount of coffee consumed. The consumption groups are labeled \texttt{A - E}.

  - \texttt{A}: 1 cup or less of caffeinated coffee consumed per week
  - \texttt{B}: 2 to 6 cups of caffeinated coffee consumed per week
  - \texttt{C}: 1 cup of caffeinated coffee consumed per day
  - \texttt{D}: 2 to 3 cups of caffeinated coffee consumed per day
  - \texttt{E}: 4 or more cups of caffeinated coffee consumed per day
  
a) Create a plot that shows the association between MET score and coffee consumption. Describe what you see.
```{r}
setwd("C:/Users/VARSHITH REDDY/Downloads")
load('coffee_exercise.Rdata')
str(coffee.exercise)
```
```{r}
library(ggplot2)
ggplot(data=coffee.exercise,mapping=aes(coffee.consumption,met,fill=met,col=met))+geom_point() +ggtitle("Associaton between Met score and Coffee Consuption")
#The plot below shows that level B is highly correlated with Met and level E has least correlation.
```



b) Conduct an analysis to determine whether the average physical activity level varies among the different levels of coffee consumption.
#Null Hypothesis - There is variation among average physical activity level among different coffee consumption groups.
#Alternative hypothesis - There is no variation in the average physical activity among different levels of coffee consumption.

#Here the p-value is less than 0.05, so we reject null hypothesis.
```{r}
summary(aov(coffee.exercise$met ~ coffee.exercise$coffee.consumption))
```


   i. Assess whether the assumptions for the analysis method are reasonably satisfied.
```{r}
#Two-sample t-test
pairwise.t.test(coffee.exercise$met, coffee.exercise$coffee.consumption, p.adj = 'none')
#Bonferroni correction
pairwise.t.test(coffee.exercise$met, coffee.exercise$coffee.consumption, p.adj = 'bonf')
```
   

    
   ii. Summarize the conclusions and comment on the generalizability of the study results.
#We have compared the variables in the coffee.exercise using paired.t.test, The average physical activity level varies among the different levels of coffee consumption.


### Problem 4. (100 points)

Problem Set 1 introduced data from a study assessing whether a relationship exists between the fluoride content in a public water supply and the dental caries experience of children with access to the supply. The file \texttt{water.Rdata} contains data from a study examining 7,257 children in 21 cities from the Flanders region in Belgium. 

The fluoride content of the public water supply in each city, measured in parts per million (ppm), is saved as the variable \texttt{fluoride}; the number of dental caries per 100 children examined is saved as the variable \texttt{caries}. The number of dental caries is calculated by summing the numbers of filled teeth, teeth with untreated dental caries, teeth requiring extraction, and missing teeth at the time of the study. 

a) Create a plot that shows the relationship between fluoride content and caries experience. Add the least squares regression line to the scatterplot.
```{r}
load('water.Rdata')
str(water)
#Creating a scatter plot for fluoride content and caries experience
plot.new()
plot(water$fluoride ~ water$caries, col = 'blue', pch= 16, main = 'Relation between Fluoride content and Caries rates', xlab = 'Caries rates per 100 children', ylab = 'Fluoride content (in ppm)')

#Calculating linear regression 
water_fit <- lm(fluoride ~ caries, data = water)
water_fit

#Adding the least squares regression line to scatter plot
plot(water_fit)

#ggplot
ggplot(data = water, mapping = aes(fluoride,caries))+geom_point()+ggtitle("Sctterplot of Fluoride and Caries in water data")
```

b) Based on the plot from part a), comment on whether the model assumptions of linearity and constant variability seem reasonable for these data.
#We can see that there is an existence of non0linear relationship between caries and fluoride levels which is based on the observation from the above scatter plots. As the regression line is not straight, we can say that the linearity is not reasonable. Change in caries does not change the proportionality to fluoride. Therefore we can say that assumption of constant variable are also not reasonable.

c) Use a residual plot to assess the model assumptions of linearity and constant variability. Comment on whether the residual plot reveals any information that was not evident from the plot from part b).
```{r}
# building a model
water_model<-lm(caries~fluoride, data=water)
# generating a residual plot 
library(handyplots)
resplot(water_model,highlight.outliers=TRUE)

#By looking at the residual plot we can say that there are many outliers.
```


Suppose the file \texttt{water\_new.Rdata} contains data from a more recent study conducted across 175 cities in Belgium (the data are simulated). Repeat the analyses from parts a) - c) with the new data.

d) Create a plot that shows the relationship between fluoride content and caries experience in the new data. Add the least squares regression line to the scatterplot.
```{r}
load('water_new.Rdata')
str(water.new)

plot(water.new$fluoride ~ water.new$caries, col = 'blue', pch = 16, main = 'Relation between fluoride content and caries rates', xlab = 'Caries rates per 100 children', ylab =  'Fluoride content (in ppm)')

#Calculating the linear regression
water_fit_new <- lm(fluoride ~ caries, data = water.new)
water_fit_new

#adding the least squares regression line to the scatter plot 
plot(water_fit_new)

#ggplot
ggplot(data = water.new, mapping=aes(fluoride,caries))+geom_point()+geom_smooth()+ggtitle('A scatterplot of Fluoride and Caries in the water data')
```

e) Based on the plot from part d), comment on whether the model assumptions of linearity and constant variability seem reasonable for these data.
#We find that there is non-linear relationship between caries and fluoride levels based on observations of scatterplot above. We can see that regression line is not linear, so the assumptions of linearity does not seem reasonable. Change in caries does not change the proportional change to fluoride. So the assumptions of constant variable are reasonable.

f) Use a residual plot to assess the model assumptions of linearity and constant variability. Comment on whether the residual plot reveals any information that was not evident from the plot from part e).
```{r}
#Showing the residual plot by below function
plot(water_fit_new)

#We also have QQplot, Scale-location, Residuals vs leverage
```
    

## Unit 7


### Problem 5. (200 points)

In Units 6 and 7, you have become familiar with the Prevention of REnal and Vascular END-stage Disease (PREVEND) study, which took place between 2003 and 2006 in the Netherlands. Clinical and demographic information for 500 individuals are stored as \texttt{prevend.samp} in the \texttt{oibiostat} package.

The PREVEND data were mainly used throughout the Unit 7 lectures to demonstrate one application of multiple regression: estimating the association between a response variable and primary predictor of interest while adjusting for confounders. Unit 7, Lab 3 discusses a model for the association of RFFT score with statin use that adjusts for age, educational level, and presence of cardiovascular disease. This question uses the PREVEND data in the context of explanatory model building.

Suppose that you have accepted a request to do some consulting work for a friend. Your task is to develop a prediction model for RFFT score based on the following possible predictor variables and the data in \texttt{prevend.samp}. 

\begin{center}
\begin{tabular}{r|l}
\textbf{Variable} & \textbf{Description} \\
\hline
\texttt{Age} & age in years \\
\texttt{Gender} & gender, coded \texttt{0} for males and \texttt{1} for females \\
\texttt{Education} & highest level of education \\
\texttt{DM} & diabetes status, coded \texttt{0} for absent and \texttt{1} for present \\
\texttt{Statin} & statin use, coded \texttt{0} for non-users and \texttt{1} for users \\
\texttt{Smoking} & smoking, coded \texttt{0} for non-smokers and \texttt{1} for smokers \\
\texttt{BMI} & body mass index, in $kg/m^2$ \\
\texttt{FRS} & Framingham risk score, measure of risk for cardiovascular event with 10 years
\end{tabular}
\end{center}

The variable \texttt{Education} is coded \texttt{0} for primary school, \texttt{1} for lower secondary education, \texttt{2} for higher secondary school, and \texttt{3} for university. A higher FRS indicates higher risk of a cardiovascular event.

Your friend has requested that your final model have no more than two predictor variables. Additionally, your friend would like you to predict the mean RFFT score for a female individual of age 55 with a university education, no diabetes, no statin use, who is not a smoker, has BMI of 24, and FRS of 5. Use only the information necessary to make a prediction from your model.

In your solution, briefly explain the work done at each step of developing the final model and evaluate the final model's strengths and weaknesses.

Please consider the following sections for your solution: 

*Data Exploration*
#Numericalsummaries
```{r}
library(oibiostat)
data("prevend.samp")
str(prevend.samp)
```


```{r}
summary(prevend.samp$Age)
summary(prevend.samp$Gender)
summary(prevend.samp$Education)
summary(prevend.samp$DM)
summary(prevend.samp$Statin)
summary(prevend.samp$Smoking)
summary(prevend.samp$BMI)
summary(prevend.samp$FRS)
```
#correlation
```{r}
cor(prevend.samp$RFFT, prevend.samp$Age)
cor(prevend.samp$RFFT, prevend.samp$Gender)
cor(prevend.samp$RFFT, prevend.samp$Education)
cor(prevend.samp$RFFT, prevend.samp$DM)
cor(prevend.samp$RFFT, prevend.samp$Statin)
cor(prevend.samp$RFFT, prevend.samp$Smoking)
cor(prevend.samp$RFFT, prevend.samp$BMI)
cor(prevend.samp$RFFT, prevend.samp$FRS)
```
#Creating Histograms

``` {r, fig.width = 8, fig.height = 6, echo = FALSE}
par(mfrow = c(2, 3))

#summarize potential predictors
hist(prevend.samp$Age, col = "blue")

hist(prevend.samp$Gender, col = "red")

hist(prevend.samp$Education, col = "orange")

hist(prevend.samp$RFFT, col = "darkmagenta")

hist(prevend.samp$DM, col = "pink")

hist(prevend.samp$Smoking, col = "purple")

hist(prevend.samp$BMI, col = "blue")

hist(prevend.samp$Statin, col = "red")

hist(prevend.samp$FRS, col = "orange")
```

#transformvariables
```{r}
prevend.samp$log.BMI = log(prevend.samp$BMI)
```



*Initial Model Fitting*
```{r}
#Including the selected variables to he model
model0 <- lm(RFFT ~ Age + Education + BMI + FRS, data = prevend.samp)
model0
#Fitting the model
model_Fit <- lm(RFFT ~ Age + Education + BMI + FRS, data = prevend.samp)
model_Fit
summary(model_Fit)
```




*Model Comparison*
```{r}
#model excluding Age
model1 = lm(RFFT ~ Education + BMI + FRS, data = prevend.samp)
summary(model1)$adj.r.squared

#model excluding Education
model1 = lm(RFFT ~ Age + BMI + FRS, data = prevend.samp)
summary(model1)$adj.r.squared

#model excluding BMI
model1 = lm(RFFT ~ Age + Education + FRS, data = prevend.samp)
summary(model1)$adj.r.squared

#model excluding FRS
model1 = lm(RFFT ~ Age + Education + BMI, data = prevend.samp)
summary(model1)$adj.r.squared
```

```{r}

#load color package
library(RColorBrewer)
boxplot(RFFT ~ Education, data = prevend.samp,
main = "Education correspondance to RFFT",
col = brewer.pal(5, "Blues"))

boxplot(RFFT ~ Age, data = prevend.samp,
main = "Age correspondance to RFFT",
col = brewer.pal(5, "Blues"))
```


```{r}
#create the Age.binary variable
prevend.samp$Age.binary = prevend.samp$Age
#redefine the levels of Age
levels(prevend.samp$Age.binary) = list(Below_sixty = c("36",
                                                       "39",
                                                       "42",
                                                       "45",
                                                       "48",
                                                       "51",
                                                       "54",
                                                       "57"),
                                   Sixty_and_above = c("60",
                                                       "63",
                                                       "66",
                                                       "69",
                                                       "72",
                                                       "75",
                                                       "78",
                                                       "81"))

#Fitting the created model with age.binary variable
model_age.binary <- lm(RFFT ~ Age.binary + Education + BMI, data = prevend.samp)
summary(model_age.binary)$adj.r.squared

#Creating Education.binary variable
prevend.samp$Education.binary = prevend.samp$Education
levels(prevend.samp$Education.binary) = list(lesser_education = c(0,1), higher_education = c(2,3))

#Fitting the created model Education.binary
model_education.binary <- lm(RFFT ~ Age.binary + Education.binary + BMI, data = prevend.samp)
summary(model_education.binary)$adj.r.squared

```
We can see that there is no improvement in the models, they have produced same scores. Therefore we are using the interaction terms for better performance of the model
```{r}
model_interaction <- lm(RFFT ~ Age.binary * Education.binary + BMI, data = prevend.samp)
summary(model_interaction)$adj.r.squared
```
Here we can see the performance of the model has improved.

*Model Assessment*
```{r}
#Creating a residual plot for the model
final_model = model_interaction

#Creating a Q-Q plot for residual
qqnorm(resid(final_model))
pch = 21
main = "Q-Q plot of Model Residuals"
       
#Adding a straight diagonal line to the plot
qqline(resid(final_model))       
```




*Conclusions*

```{r}

#Final model summary
summary(final_model)
```



## Unit 8

### Problem 6. (200 points)

Biological ornamentation refers to features that are primarily decorative, such as the elaborate tail feathers of a peacock. The evolution of ornamentation in males has been extensively researched; there are many studies exploring how male ornamentation functions as a signal of phenotypic and/or genetic quality to potential mates. In contrast, there are few studies investigating female ornamentation.\footnote{Freeman-Gallant, et al., J Evol. Biol. (2014) 27: 982-991 \url{doi: 10.1111/jeb.12369}.}

Some biologists have hypothesized that there is strong natural selection against overly conspicuous female ornaments. Bright or colorful plumage in females might be expected to increase the incidence of predation on nests for species in which females incubate eggs. Female ornamentation might also undergo positive selection, functioning in sexual signaling like male ornamentation, and indicating desirable qualities such as high immune function. 

The data in the file \texttt{rubythroats.Rdata} are from a study of 83 female rubythroats, a bird species in which both males and females exhibit a brightly colored red patch on the throat and breast (referred to as a "bib"). In rubythroats, females incubate the eggs, while males provide food to females to facilitate uninterrupted incubation. 

  - \texttt{survival}: records whether the bird survived to return to the nesting site the subsequent year, \texttt{yes} if the female was observed and \texttt{no} if the female was not observed
  
  - \texttt{weight}: weight of the bird, measured in grams
  
  - \texttt{wing.length}: wing length of the bird, measured in millimeters
  
  - \texttt{tarsus.length}: tarsus (i.e., leg) length of the bird, measured in millimeters
  
  - \texttt{first.clutch.size}: number of eggs in the first clutch laid during the first year that the bird was observed
  
  - \texttt{nestling.fate}: whether the nestlings from the first clutch survived to fledging (\texttt{Fledged}) or were lost to predation (\texttt{Predated})
  
  - \texttt{second.clutch}: whether the bird laid a second clutch during the first year that the bird was observed, recorded as \texttt{Yes} for laying a second clutch and \texttt{No} for otherwise
  
  - \texttt{carotenoid.chroma}: a measure of the abundance of red carotenoid pigment in feathers, as measured from a sample of four feathers taken from the center of the bird's bib. Larger numbers indicate higher levels of pigment in the feathers and a more saturated red color.
  
  - \texttt{bib.area}: the total area of the bird's bib, measured in millimeters squared
  
  - \texttt{total.brightness}: a measure of bib brightness, calculated from spectrometer analyses. Larger numbers indicate a brighter red color.
  
You will be conducting an analysis of the results in order to investigate how bib attributes and other phenotypic characteristics of female birds are associated with measures of fitness.
```{r}
#Loading the data
setwd("C:/Users/VARSHITH REDDY/Downloads")
load('rubythroats.Rdata')
str(rubythroats)
```

a) Fit a model to predict nestling fate from female bib characteristics (carotenoid chroma, bib area, total brightness) and female body characteristics (weight, wing length, tarsus length). Identify the slope coefficients significant at $\alpha = 0.10$, and provide an interpretation of these coefficients in the context of the data.
# The slope coeeficients of predictors Total brightness with p value = 0.0273 and Wing length with p value = 0.02627 are significant at alpha 0.10. When there is increase in total brightness, the estimated log odds of nestling fate is reduced by 0.130880 when other variables are constant. Similarly the wing length is increased , the estimated log odds of nestling fate is increased by 0.521821. This provides an interpretation that total brightness and wing length are the most important and necessary predictors for the residual value. 
```{r}
rubythroats$nestling.fate <- as.factor(rubythroats$nestling.fate)
log_reg_fit <- glm(nestling.fate ~ carotenoid.chroma + bib.area + total.brightness + weight + wing.length + tarsus.length, family = binomial(link = 'logit'),data = rubythroats)
log_reg_fit
summary(log_reg_fit)
```


b) Investigate the factors associated with whether a female lays a second clutch during the first year that she was observed. 

   i. Is there evidence of a significant association between nestling fate and whether a female lays a second clutch? If so, report the direction of association.
#As the p vale is 5.499e-05 which is less than 0.05, we reject null hypothesis. And there is an association. The slope of the interaction term indicates the association between nestling fate and second clutch to be positive as the value is 3.4045 and the p-value is less than 0.05. The proportion of birds predated  when they did not lay second clutch is 0.6825397 and the proportion of birds fledged when they did not lay second clutch is 0.3174603    
```{r}
table_ns <- table(rubythroats$nestling.fate, rubythroats$second.clutch)
table_ns

#Describing Null hypothesis - no association between nestling fate and second clutch
#Describing alternative hypithesis - there is association betwee nestling fate and second clutch 
chisq.test(rubythroats$nestling.fate, rubythroats$second.clutch)

#Testing of Interaction
nestling_secondeggs <- glm(nestling.fate~second.clutch, family = binomial(link = 'logit'), data = rubythroats)

nestling_secondeggs
summary(nestling_secondeggs)
```

    
  ii. Fit a model to predict whether a female lays a second clutch from nestling fate and bib characteristics. Identify the two predictors that are most statistically significantly associated with the response variable.
    
#Answer - The two predictors most statistically significantly associated with laying a second clutch are total brightness ($p = 0.030$) and nestling fate ($p = 0.0015$). 
```{r}
#fitting new model
#predictors - nestlingfate, carotid chroma, bib area, total brightness
#response variable - second clutch
#second clutch as factor
rubythroats$second.clutch <- as.factor(rubythroats$second.clutch)
new_model_ruby <- glm(second.clutch ~ nestling.fate + carotenoid.chroma + bib.area + total.brightness, family = binomial(link = 'logit'), data = rubythroats)
summary(new_model_ruby)
```

    

 iii. Fit a new model to predict whether a female lays a second clutch using the two predictors identified in part ii. and their interaction. Interpret the model coefficients in the context of the data.
#The coefficients for nestling.fate is 0.0103 and total brightness is 0.16532 . The p value of the model is 0.00136 signifying that both the predictor variables influence the model. Nestling fate shows positive correlation while total brightness has no correlation with second clutch. The individual p value of total brightness shows that it is not significantly associated with second clutch.
```{r}
model_secondclutch <- glm(second.clutch ~ nestling.fate + total.brightness, family = binomial(link = 'logit'), data = rubythroats)
summary(model_secondclutch)
```

        

c) Investigate the factors associated with whether a female survives to return to the nesting site the subsequent year.

   i. Fit a model to predict survival from bib characteristics, female body characteristics, first clutch size, and whether a second clutch was laid. Identify factors that are positively associated with survival for the observed birds.

#The factors that are positively associated with survival of these birds include,
#weight - 0.605506
#wing length - 0.842892
#first clutch - 2.620596
#second clutch size - 1.664275

```{r}
#assigning survival as factor 
rubythroats$survival <- as.factor(rubythroats$survival)

#predicting the model
model_female_survives <- glm(survival ~ carotenoid.chroma + bib.area + total.brightness + weight +wing.length + tarsus.length + first.clutch.size + second.clutch, family = binomial(link = 'logit'), data = rubythroats)
summary(model_female_survives)
```


   
   ii. Fit a new model with only the significant predictors from the previous model; let $\alpha = 0.10$. Comment on whether this model is preferable to the one fit in part i.
#Here the p value is 0.05 and the one in part 1 is 0.2852. So, this model predicts better than that in model 1. Also the predictors in this model have significance levels less than alpha contributing to the prediction better. 
```{r}
#model fitting
model_significant <- glm(survival ~ carotenoid.chroma + total.brightness + wing.length + first.clutch.size, family = binomial(link = 'logit'), data = rubythroats)
summary(model_significant)
```
```{r}
#Better parsimonious model of the ones fit in parts i. and ii. 
AIC(model_female_survives)
AIC(model_significant)
```

 
   *For parts iii. and iv., use the better parsimonious model of the ones fit in parts i. and ii.*
 
  iii. Compare the odds of survival for a female who laid 5 eggs in her first clutch to the odds of survival for a female who laid 3 eggs in her first clutch, if the females are physically identical and both laid a second clutch.

#We use the better parsimonious model, which is model_significant.
#The female who laid five eggs in the first clutch has odds of survival 3.901085
#The female who laid five eggs in the first clutch has odds of survival 0.1708379

#Female who laid five eggs in the first clutch has 22.83501 times survival to female who laid three eggs in the first clutch .

```{r}
#First clutch size 5

log.odds.eggs_5 = predict(model_significant, data.frame(carotenoid.chroma = 0.991766, total.brightness = 17.25410, wing.length = 53.0, first.clutch.size = 5))

exp(log.odds.eggs_5)

#First clutch size 3
log.odds.eggs_3 = predict(model_significant, data.frame(carotenoid.chroma = 0.99176, total.brightness = 17.25410, wing.length = 53.0, first.clutch.size = 3))

exp(log.odds.eggs_3)

#comparing the odds of survival between 2 females 
exp(log.odds.eggs_5/exp(log.odds.eggs_3))
```

  
   iv. Suppose female A has bib area 350 $mm^2$, total brightness of 35, carotenoid chroma 0.90, tarsus length of 19.5 $mm$, wing length 51 $mm$, weighs 10.8 $g$, lays 4 eggs in her first clutch, and lays a second clutch. Female B has bib area 300 $mm^2$, total brightness of 20, carotenoid chroma 0.85, tarsus length of 19.0 $mm$, wing length 50 $mm$, weighs 10.9 $g$, lays 3 eggs in her first clutch, and lays a second clutch. Compare the odds of survival for females A and B.
#The odds of survival for Female A = 0.2047139
#The odds of survival for Female B = 0.1341531
#The odds of survival of female A is 1.525972 time greater tham female B 

```{r}
#Female A odds of survival

p_1 = predict(model_significant, newdata = data.frame(carotenoid.chroma = 0.90, total.brightness = 35, wing.length=51, first.clutch.size = 4))
exp(p_1)

#Female B odds of survival 

p_2 = predict(model_significant, newdata = data.frame(carotenoid.chroma = 0.85, total.brightness=20, wing.length=50, first.clutch.size =3))
exp(p_2)

#Now comapring the odds of survival between female A and B 

exp(p_1)/exp(p_2)
```

      
    
d) Biological fitness refers to how successful an organism is at surviving and reproducing. Based on the results of your analysis, briefly discuss whether female ornamentation seems beneficial for fitness in this bird species. Limit your response to at most ten sentences. You do not need to reference specific numerical results/models from the analysis.

#Biological fitness plays a vital role for an organism to survive at its best.
#It maintains the perfect lifecycle of each and every organism to survive and reproduce.
#We have done the analysis on rubythroat datset which fits the above sentence perfectly.
#After the analysis we have done on different variables which affects the organisms life, we can say that biological fitnesss is the most important factor for an organism to live in the world of "Survival of fitness.
#For example to answer the reproduction of the female we have take the variables such as carotenoid.chroma, total.brightness, wing.length and fisrst.clutch size. After analyzing all these variables above we can conclude that odds of survival and reproduction of the birds are great when they are biologically fit. It applies to all the parameters above. 


